\section{System architecture and data flow}
\label{sec:architecture}

This section is the map for the rest of the document. It introduces the four tiers and their
responsibilities, the two operating modes, and the four end-to-end data flows that together cover
everything the system does: live monitoring, training capture, device onboarding, and model delivery.

\subsection{Topology}

Meowtion is a four-tier system, \cref{fig:architecture}: the \textbf{collar} on the cat, the
\textbf{station} in the home, the \textbf{cloud} backend, and the owner's \textbf{dashboard}. The
defining constraint is that the collar is BLE-only and battery-powered, with no route to the internet
of its own. The mains-powered station bridges that gap, giving the collar a cheap, always-on
Bluetooth peer and a Wi-Fi path to the cloud. Everything above the collar is conventional web
infrastructure; everything at the collar is shaped by power and radio budget.

\begin{figure}[ht]
  \centering
  \resizebox{\textwidth}{!}{%
  \begin{tikzpicture}[node distance=9mm and 13mm]
    \node[mwnode] (cat) {Cat};
    \node[mwnode, right=of cat, fill=mwlav!12, draw=mwlav] (collar)
      {\textbf{Collar}\\\footnotesize XIAO nRF52840 Sense\\\footnotesize Zephyr\,/\,NCS, battery\\\footnotesize sensing + on-device AI};
    \node[mwnode, right=of collar, fill=mwlav!12, draw=mwlav] (station)
      {\textbf{Station}\\\footnotesize XIAO ESP32-S3\\\footnotesize ESP-IDF, mains\\\footnotesize BLE central + Wi-Fi gateway};
    \node[mwnode, right=of station] (cloud)
      {\textbf{Firebase}\\\footnotesize Auth, Realtime DB\\\footnotesize Storage, Functions};
    \node[mwnode, below=15mm of station] (dash)
      {\textbf{Dashboard}\\\footnotesize Streamlit + web SDK};
    \node[mwnode, left=of dash] (owner) {Owner};

    \draw[mwedge] (cat) -- node[above,font=\scriptsize,text=mwmut]{sensing} (collar);
    \draw[mwedge] (collar) -- node[above,font=\scriptsize,text=mwmut]{BLE} (station);
    \draw[mwedge] (station) -- node[above,font=\scriptsize,text=mwmut]{Wi-Fi\,/\,HTTPS} (cloud);
    \draw[mwedge] (cloud.south) |- (dash.east);
    \draw[mwedge] (dash) -- (owner);
  \end{tikzpicture}%
  }
  \caption{End-to-end topology. The collar is BLE-only, so the always-on station is its gateway to
           the cloud. The dashboard is a separate client of the same backend.}
  \label{fig:architecture}
\end{figure}

\subsection{Component responsibilities}

\Cref{tab:responsibilities} summarises what each tier is responsible for.

\begin{center}
  \captionof{table}{What each tier is responsible for.}
  \label{tab:responsibilities}
  \begin{tabularx}{\textwidth}{@{}lX@{}}
    \toprule
    \textbf{Tier} & \textbf{Responsibility} \\
    \midrule
    Collar    & Sense (IMU + microphone), classify behaviour on-device, advertise the result as BLE
                telemetry, and, when a station subscribes, stream paired audio\,+\,IMU clips for
                training and accept models over the air. Holds no credentials. \\
    Station   & Scan for and relay registered collars to the cloud; capture and upload training
                clips; fetch trained models from Storage and push them to the collar over BLE.
                Authenticates to the cloud with a per-device token, not a login. \\
    Cloud     & Firebase: authenticate owners, store the live data and history (Realtime Database),
                store training clips and trained models (Storage), and run server-side logic (Cloud
                Functions for training and authenticated clip upload). Enforces per-owner isolation. \\
    Dashboard & The owner-facing web app: a live current-state view and longitudinal health
                analytics, plus a static front door for sign-in and device registration and a
                dev console for data capture and training. \\
    \bottomrule
  \end{tabularx}
\end{center}

\subsection{Operating modes}

A station runs in one of two modes at a time, selected per station from the dashboard. The mode
governs what crosses the BLE link and what the station does with it.

\begin{description}[leftmargin=1.6em,style=nextline]
  \item[Production mode] The everyday monitoring mode and privacy-preserving default. The collar
        classifies on-device and advertises only the result; the station reads it and relays it to
        the cloud. No audio or raw motion leaves the collar.
  \item[Training mode] The data-collection mode for teaching a new behaviour. While a registered
        collar is in range, the station subscribes to the collar's audio service; the collar streams
        paired audio\,+\,IMU clips, which the station slices and uploads and the owner labels and
        trims in the dev console. It is used deliberately and temporarily, not continuously.
\end{description}

\subsection{End-to-end data flows}

Four flows fully describe the system's behaviour. Each is given below as a concrete sequence of steps
across the tiers; the wire formats are specified in \cref{sec:protocols} and the security rules in
\cref{sec:security-privacy}.

\paragraph{Flow 1 --- Live monitoring (production).}
The steady state of a deployed system.
\begin{enumerate}[leftmargin=1.6em,itemsep=2pt]
  \item The collar samples its IMU continuously and runs the on-device cascade
        (\cref{sec:on-device-ai}); when the motion result is ambiguous it briefly samples the
        microphone and runs the audio model to confirm.
  \item The collar encodes the outcome (class index, confidence, step count, battery) into an 8-byte
        BLE telemetry advertisement (\cref{sec:protocols}). Its BLE address identifies the collar, so
        the payload carries no identifier.
  \item The station, scanning continuously, hears the advertisement from a \emph{registered} collar,
        timestamps it, maps the class index to the owner's label set, and writes the current state
        plus an event record to the Realtime Database over authenticated HTTPS.
  \item The dashboard, reading the same database, refreshes the live card within seconds and folds
        the new event into the history and health analytics (\cref{sec:dashboard}).
\end{enumerate}

\begin{figure}[ht]\centering
  \resizebox{\textwidth}{!}{%
  \begin{tikzpicture}[node distance=18mm, every node/.style={font=\footnotesize}]
    \node[mwnode] (collar) {Collar\\\scriptsize classify on-device};
    \node[mwnode, right=of collar] (station) {Station\\\scriptsize map class\,$\to$\,label};
    \node[mwnode, fill=mwlav!8, right=of station] (rtdb) {Realtime DB\\\scriptsize current + events};
    \node[mwnode, right=of rtdb] (dash) {Dashboard\\\scriptsize live + history};
    \draw[mwedge] (collar) -- node[above,text=mwmut,font=\scriptsize]{telemetry AD (BLE)} (station);
    \draw[mwedge] (station) -- node[above,text=mwmut,font=\scriptsize]{HTTPS (token)} (rtdb);
    \draw[mwedge] (rtdb) -- node[above,text=mwmut,font=\scriptsize]{read} (dash);
  \end{tikzpicture}}
  \caption{Flow~1, live monitoring. The collar advertises an 8-byte telemetry packet; the station
           maps the class index to a label and writes the live state and events to the database, which
           the dashboard reads.}
  \label{fig:live-flow}
\end{figure}

\paragraph{Flow 2 --- Training capture.}
How labelled data is collected to teach a behaviour.
\begin{enumerate}[leftmargin=1.6em,itemsep=2pt]
  \item In the dev console the owner puts a station into training mode and selects the action being
        recorded (the ``live label'').
  \item The station connects to the in-range collar and subscribes to its audio characteristic. The
        collar streams continuously in 100\,ms frames, each a header followed by an audio payload and
        a time-aligned IMU payload (\cref{sec:protocols}).
  \item The station concatenates frames, slices them into fixed-length clips, and POSTs each clip's
        bytes to the \texttt{upload\_clip} Cloud Function with its device token. The function
        verifies the token's owner and writes the clip to that owner's Storage area; the station
        records the clip's metadata (label, paths, IMU presence) in the database.
  \item The owner reviews clips in the dev console: trims the waveform, confirms or changes the label,
        and discards bad takes. The labelled clips are the training set for Flow~4.
\end{enumerate}

\paragraph{Flow 3 --- Device onboarding.}
How a station and collar are registered to an owner, performed once per device.
\begin{enumerate}[leftmargin=1.6em,itemsep=2pt]
  \item The owner signs in on the static front door and connects the station over USB using the
        browser's Web Serial API. The page mints a random per-device \emph{token}, records it under
        the owner's account (\texttt{deviceTokens/<token>/owner}) and as a station device entry, and
        hands the station its Wi-Fi credentials, the owner id, and the token over the serial link.
        The station stores these in non-volatile storage; the token is its sole credential.
  \item The station comes online, syncs time, and begins scanning, publishing any unregistered
        collars it hears to its \texttt{seen} list.
  \item The owner registers a heard collar with one tap in the dashboard. The collar then appears as a
        device on the account and is relayed by the station. Deleting the token revokes the device.
\end{enumerate}

\begin{figure}[ht]\centering
  \begin{tikzpicture}[node distance=9mm, every node/.style={font=\footnotesize}]
    \node[mwterm] (a) {Owner signs in};
    \node[mwflow, below=of a, text width=7.2cm, align=left] (b)
      {Connect the station over USB (Web Serial). The dashboard mints a random token, records
       \texttt{deviceTokens/\allowbreak <token>/owner} and a station device entry};
    \node[mwflow, below=of b, text width=7.2cm, align=left] (c)
      {Provision the station over USB: Wi-Fi credentials, the owner id, and the token (stored in NVS)};
    \node[mwflow, below=of c, text width=7.2cm, align=left] (d)
      {Station comes online, syncs time, scans, and publishes unregistered collars it hears (\texttt{seen})};
    \node[mwterm, below=of d] (e) {Owner registers a heard collar (one tap)};
    \draw[mwedge] (a) -- (b); \draw[mwedge] (b) -- (c); \draw[mwedge] (c) -- (d); \draw[mwedge] (d) -- (e);
  \end{tikzpicture}
  \caption{Flow~3, connecting devices. A station is paired over USB and provisioned with a scoped
           token; once online it surfaces nearby collars, which the owner registers with one tap.}
  \label{fig:onboarding}
\end{figure}

\paragraph{Flow 4 --- Model training and over-the-air delivery.}
How labelled clips become a model running on the collar.
\begin{enumerate}[leftmargin=1.6em,itemsep=2pt]
  \item The owner triggers training from the dev console. The \texttt{train} Cloud Function reads the
        account's labelled clips, derives the class set from the labels present, trains an int8 IMU
        model and an int8 audio model, uploads them to Storage under \texttt{models/<uid>/}, and bumps
        the account's model version (\cref{sec:training-pipeline}).
  \item The station learns a new version is available (mirrored into its own device config), downloads
        each model from Storage, and, over a brief dedicated BLE connection, streams it to the collar
        with a CRC-checked begin/data/end protocol (\cref{sec:protocols}).
  \item The collar stages the model into its dedicated flash partition, verifies the CRC, loads it,
        and switches from advertising simulated state to advertising real classifier output. The path
        requires no firmware reflash.
\end{enumerate}

\subsection{Multi-tenant data model}

Meowtion is multi-tenant. Each owner creates an account and registers their own station(s) and
collar(s); all of an owner's data lives under \texttt{users/<uid>/} in the Realtime Database, and the
database and storage rules scope every read and write to the owning account
(\cref{sec:cloud-backend,sec:security-privacy}). The backend enforces isolation, not the client.
Battery devices never hold a login: a station authenticates with its scoped, revocable token, and a
collar holds nothing at all (the station relays it). A single public, read-only demo account,
designated in server-side config, lets anyone explore a populated dashboard without signing up.

\subsection{Technology summary}

\Cref{tab:stack} lists the technology used at each tier.

\begin{center}
  \captionof{table}{Technology at each tier.}
  \label{tab:stack}
  \begin{tabularx}{\textwidth}{@{}lX@{}}
    \toprule
    \textbf{Tier} & \textbf{Stack} \\
    \midrule
    Collar    & Seeed XIAO nRF52840 Sense; Zephyr / nRF Connect SDK (v3.3.1); TensorFlow Lite Micro
                with CMSIS-NN kernels \\
    Station   & Seeed XIAO ESP32-S3; ESP-IDF; NimBLE (BLE central) + Wi-Fi station \\
    Cloud     & Firebase: Authentication, Realtime Database, Cloud Storage, Python Cloud Functions
                (2nd gen) \\
    Dashboard & Streamlit (Python) + Firebase web SDK (JavaScript); hosted on Streamlit Community
                Cloud \\
    Training  & TensorFlow / Keras, executed server-side in a Cloud Function; int8 post-training
                quantisation to TensorFlow Lite \\
    \bottomrule
  \end{tabularx}
\end{center}
