\section{Station firmware}
\label{sec:firmware-station}

The station runs on ESP-IDF~\cite{espidf} on the XIAO ESP32-S3. It is a BLE central that hears registered collars
and a Wi-Fi gateway that relays them to the owner's Firebase account. It also captures short
audio\,+\,IMU clips for training and uploads them, and it delivers trained models to the collar over
BLE. It has no Firebase login: Wi-Fi credentials, the owner id, and a per-device token are
provisioned over USB serial and stored in non-volatile storage (NVS), and the database authorises the
station because that token is registered to the owner.

\subsection{Source layout}

\begin{table}[ht]
  \centering
  \caption{Station firmware modules (\texttt{firmware/station/main/}).}
  \label{tab:station-src}
  \begin{tabularx}{\textwidth}{@{}lX@{}}
    \toprule
    \textbf{File} & \textbf{Responsibility} \\
    \midrule
    \texttt{main.c}    & Boot and the roughly one-second orchestration loop \\
    \texttt{config.c}  & Provisioned credentials, NVS storage, USB-serial provisioning \\
    \texttt{board.c}   & Power-source detection (USB vs battery) \\
    \texttt{wifi.c}    & Wi-Fi station bring-up and time synchronisation \\
    \texttt{cloud.c}   & The Firebase Realtime Database / HTTP layer (the only HTTP in the firmware) \\
    \texttt{weather.c} & Local weather from the Open-Meteo feed \\
    \texttt{ble.c}     & The whole NimBLE collar subsystem (scan, relay, audio capture) behind \texttt{ble\_service()} \\
    \texttt{ota.c}     & OTA model delivery: pushes trained \texttt{.tflite} models to the collar over BLE \\
    \bottomrule
  \end{tabularx}
\end{table}

\subsection{Provisioning}

Provisioning is a single JSON line sent over USB serial; the dashboard's ``Connect a device'' form
does this automatically over Web Serial, but it is plain text and can be sent by hand for debugging:

\begin{lstlisting}[caption={Provisioning line (USB serial).},label={lst:provision}]
{"ssid":"..","pass":"..","owner":"<uid>","token":"<random>","name":"..","lat":50.8,"lon":-0.1}
\end{lstlisting}

The station stores these in NVS and can be re-provisioned at any time from the dashboard, or wiped
with \texttt{idf.py erase-flash}. The token is the station's only credential; deleting it from the
owner's account revokes the device immediately. Repository secrets such as \texttt{secrets.h} are
excluded by \texttt{.gitignore}, so no live credential is tracked.

\subsection{Relay and clip upload}

In production mode the station scans continuously, hears the telemetry advertisements of registered
collars, maps the compact result (mode, class index, confidence) onto the cat's record, and writes
both a live \texttt{current} state and discrete activity events as episodes begin and end
(\texttt{cloud.c}). Events carry the model version and class index rather than a resolved name, so the
dashboard can map them to labels even as the label set changes (\cref{sec:cloud-backend}). In training
mode it connects to an in-range collar, captures clips, and POSTs each clip's bytes to the
\texttt{upload\_clip} Cloud Function with its device token; the function verifies the token's owner and
writes the file to that owner's Storage area, since clients cannot write Storage directly. A manual
capture path ignores the signal-strength gate, used to collect behaviours such as purring that happen
while the cat rests rather than at the bowl.

\subsection{Over-the-air model delivery}

The station is the delivery vehicle for trained models, since the collar is BLE-only and never touches
Storage. The wire protocol is fixed in the shared header \texttt{firmware/common/meow\_ota.h}: the
collar is the GATT server that stages a \texttt{.tflite} into flash, the station is the client that
pushes it (\cref{sec:protocols}).

\begin{description}[leftmargin=1.6em,style=nextline]
  \item[Version gate.] The \texttt{train} function bumps \texttt{users/<uid>/models/version} and
        mirrors it into this station's own config as \texttt{modelVer} (the station may read only its
        own \texttt{users/<uid>/devices/<token>/} subtree, not \texttt{users/<uid>/models}). The
        station keeps its last-pushed version in NVS; a push is due when the config version is higher.
  \item[Download.] For each slot it fetches the model from Storage at the public-read path
        \texttt{models/<uid>/\allowbreak imu\_model.tflite} (IMU slot) or
        \texttt{models/<uid>/\allowbreak audio\_model.tflite} (audio slot). A 404 means that slot has no model
        yet and is skipped.
  \item[Push.] It computes a CRC-32/IEEE over the bytes, then over a brief dedicated BLE connection
        writes \texttt{BEGIN}, streams the model to the data characteristic in
        write-with-response chunks (the chunk write-response is the flash-write flow control), and
        writes \texttt{END}. On the collar's \texttt{OK} it records the new version in NVS; on any
        error it leaves NVS unchanged and retries later.
\end{description}

A push runs only when a registered collar is in range and audio capture is idle, so model delivery
never contends with the capture or relay path for the radio.

\subsection{Robustness: scanning after an update}

The station drives one BLE radio for several jobs, and the OTA path exposed a subtle defect. To push a
model the station must cancel scanning and open a connection; the OTA path used its own GAP callback
and originally never restarted scanning when it finished, so after an update the station went deaf and
the dashboard reported the collar offline. The fix is an idempotent \texttt{ble\_resume\_scan()},
guarded on the discovery-active flag, that the OTA path calls on both completion and bail-out. It was
verified on hardware.

\subsection{Build and flash}

The station builds with \texttt{idf.py build} and flashes over its USB serial port with
\texttt{idf.py -p <COM> flash monitor} (the repository's VS Code task \textbf{Build+Flash Station}
wraps this). It is verified building under ESP-IDF.
