\section{On-device AI}
\label{sec:on-device-ai}

All classification happens on the collar in real time, using TensorFlow Lite
Micro~\cite{david2021tflm} with the CMSIS-NN kernels~\cite{lai2018cmsisnn} for the nRF52840's
Cortex-M4F. The models are tiny int8 networks in the spirit of TinyML~\cite{warden2019tinyml}: small
enough to fit a microcontroller's RAM and run within the collar's power budget, yet accurate enough to
separate everyday behaviours.

\subsection{Three-tier cascade}

Inference cost rises sharply across the cascade, so Meowtion gates each tier behind a cheaper one and
pays for the next only when the cheaper tier cannot answer. Three tiers run in front of every reported
behaviour:

\begin{enumerate}[leftmargin=1.6em,itemsep=2pt]
  \item \textbf{Rules-based activity gate} (no model). A cheap motion-energy rule on the raw IMU
        stream decides \emph{active} vs \emph{resting}. Most of a cat's day is rest, and classifying
        stillness 50 times a second drains the cell for no information, so sustained low motion drops
        the collar into low-power rest: the 104\,Hz sampling and inference stop, the IMU falls back to
        a low-power motion watch, and the radio slows. Motion ends rest, and the whole dormant span is
        reported as a single \emph{rest} episode (\cref{sec:firmware-collar}). This gate is the
        dominant battery lever (\cref{sec:evaluation}).
  \item \textbf{IMU model} (model-based). While the cat is active, the int8 motion model runs on every
        window. It is cheap and carries most of the classification load.
  \item \textbf{Audio model} (model-based). When the IMU model is \emph{unsure} (its top class falls
        below a confidence threshold of 0.75), the short audio model is invoked to confirm. Audio is
        therefore a deployed disambiguation signal, not merely a training aid.
\end{enumerate}

The audio tier earns its keep on one hard case: eating and drinking look almost identical to a motion
sensor (head down at a bowl, similar gross motion) but sound completely different. The cascade keeps
average power low at both ends , the rules-based gate suppresses all sensing and inference while the
cat rests, and the audio path runs only to break a tie , while still recovering accuracy that motion
alone cannot provide. The policy lives in plain C, independent of the inference backend: the
rules-based gate in \texttt{activity.c}, and the model-tier gating in \texttt{classifier.c}.
\Cref{fig:cascade} shows the decision flow.

\begin{figure}[ht]
  \centering
  \begin{tikzpicture}[node distance=8mm and 14mm]
    \node[mwterm] (win) {IMU stream};
    \node[mwdecision, below=of win] (act) {moving?};
    \node[mwflow, left=of act, text width=2.3cm] (rest) {Low-power rest\\\scriptsize rest event on wake};
    \node[mwflow, below=of act] (imu) {IMU model};
    \node[mwdecision, below=of imu] (q) {top class\\confident?};
    \node[mwterm, below=of q] (out1) {Report class\\+ confidence};
    \node[mwflow, right=of q] (aud) {Audio model};
    \node[mwterm, below=of aud] (out2) {Report\\confirmed class};
    \draw[mwedge] (win) -- (act);
    \draw[mwedge] (act) -- node[above,font=\scriptsize,text=mwmut]{no} (rest);
    \draw[mwedge] (act) -- node[right,font=\scriptsize,text=mwmut]{yes} (imu);
    \draw[mwedge] (rest.north) to[bend left=40] node[above,font=\scriptsize,text=mwmut]{motion} (win.west);
    \draw[mwedge] (imu) -- (q);
    \draw[mwedge] (q) -- node[left,font=\scriptsize,text=mwmut]{yes} (out1);
    \draw[mwedge] (q) -- node[above,font=\scriptsize,text=mwmut]{no} (aud);
    \draw[mwedge] (aud) -- (out2);
  \end{tikzpicture}
  \caption{The three-tier cascade. The rules-based activity gate runs first: while the cat is still
           the collar rests in low power and the dormant span is logged as one rest event; on motion
           the IMU model runs on every window, and the audio model is invoked only when the IMU
           model's top class is below the confidence threshold. With no model loaded the model tiers
           return \texttt{UNKNOWN}.}
  \label{fig:cascade}
\end{figure}

\subsection{Matching training and inference}

The on-device input transform must mirror the training transform exactly; this is the single most
important correctness property. A model trained on one representation and fed another at inference
produces confident nonsense. Meowtion enforces the match on both paths:

\begin{itemize}[leftmargin=1.5em]
  \item \textbf{Motion.} The rolling window is normalised per channel (subtract each channel's mean
        over the window, divide by the global maximum absolute value with a small epsilon) and
        quantised to int8 using the input tensor's own scale and zero-point.
  \item \textbf{Audio.} Inference audio is decimated and encoded to match the training
        representation: 8\,kHz $\mu$-law, produced by the same conversion used to prepare the training
        clips (\cref{sec:protocols}). A mismatch here was an early source of confident
        misclassification.
\end{itemize}

To avoid a large floating-point scratch buffer on a RAM-constrained part, the IMU path quantises
directly from the int16 source in a few passes (per-channel mean, then global max-abs, then write
int8) rather than materialising a float window (\cref{lst:quant}).

\begin{lstlisting}[language=C,caption={Direct int16-to-int8 quantisation, mirroring training (sketch).},label={lst:quant}]
// pass 1: per-channel mean over the window
// pass 2: global max-abs of (sample - channel_mean)
// pass 3: int8 = round( x_norm / scale ) + zero_point,
//         where x_norm = (sample - channel_mean) / (maxabs + 1e-6)
// scale and zero_point are read from the input tensor itself.
\end{lstlisting}

\subsection{Model architecture}

Each model stage is a small 1-D convolutional network trained in the cloud and exported as an int8
\texttt{.tflite} (\cref{sec:training-pipeline}). The IMU stage runs over a rolling window of motion
samples (six channels: accelerometer and gyroscope x/y/z); the audio stage runs over one second of
8\,kHz audio (8000 samples).

Both stages are the small 1-D CNNs defined in the training pipeline (\cref{sec:training-pipeline}).
Every convolution uses \texttt{same} padding and ReLU; with a stride $s$ the output length is
$\lceil L/s \rceil$. The classifier head has $N$ outputs, where $N$ is the number of action classes
discovered from the labelled data, so the head and the totals scale with $N$. The exported int8
model sizes are reported in \cref{tab:footprint}.

\begin{table}[ht]
  \centering
  \caption{Per-stage network architecture. \texttt{k}\,$=$\,kernel size, \texttt{s}\,$=$\,stride;
           $N$ is the action-class count (discovered from the data). Parameter counts are for the
           trained float model; export quantises weights to int8.}
  \label{tab:arch}
  \small
  \begin{tabular}{@{}lllr@{}}
    \toprule
    \textbf{Stage} & \textbf{Layer} & \textbf{Output} & \textbf{Params} \\
    \midrule
    IMU
      & Input                          & $104\times6$  & 0 \\
      & Conv1D 16, k5, s1, ReLU        & $104\times16$ & 496 \\
      & MaxPool1D 2                    & $52\times16$  & 0 \\
      & Conv1D 32, k5, s1, ReLU        & $52\times32$  & 2\,592 \\
      & MaxPool1D 2                    & $26\times32$  & 0 \\
      & Conv1D 64, k3, s1, ReLU        & $26\times64$  & 6\,208 \\
      & GlobalAvgPool1D                & $64$          & 0 \\
      & Dense 32, ReLU (Dropout 0.3)   & $32$          & 2\,080 \\
      & Dense $N$, Softmax             & $N$           & $33N$ \\
    \cmidrule(l){2-4}
      & \textbf{Total}                 &               & $11\,376 + 33N$ \\
    \midrule
    Audio
      & Input                          & $8000\times1$ & 0 \\
      & Conv1D 8, k9, s4, ReLU         & $2000\times8$ & 80 \\
      & Conv1D 16, k9, s4, ReLU        & $500\times16$ & 1\,168 \\
      & Conv1D 32, k5, s2, ReLU        & $250\times32$ & 2\,592 \\
      & Conv1D 32, k3, s2, ReLU        & $125\times32$ & 3\,104 \\
      & GlobalAvgPool1D                & $32$          & 0 \\
      & Dense 32, ReLU (Dropout 0.3)   & $32$          & 1\,056 \\
      & Dense $N$, Softmax             & $N$           & $33N$ \\
    \cmidrule(l){2-4}
      & \textbf{Total}                 &               & $8\,000 + 33N$ \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Runtime models and model-free safety}

No model is compiled into the firmware. Each stage loads its \texttt{.tflite} at runtime
(\texttt{clf\_set\_imu\_model} / \texttt{clf\_set\_audio\_model}) from the dedicated flash partition,
delivered over the air (\cref{sec:firmware-collar,sec:protocols}), so a re-trained model swaps in
without reflashing. The engine is also safe with no model present: the presence queries return false
and the cascade returns \texttt{UNKNOWN} rather than a wrong answer. This is the state of a freshly
flashed collar before its first model arrives.

\subsection{Memory budget}

TensorFlow Lite Micro uses a fixed, statically allocated tensor arena (this firmware uses no heap).
The two stages never infer at once , the cascade runs the motion model first and consults audio only
when unsure , so rather than reserve an arena each, they share a \emph{single} arena: the collar loads
exactly one model into it at a time and unloads that model before arming the other, so only one
model's tensors are ever resident. The arena is sized to the larger (audio) stage; the smaller motion
model fits inside it. The value in \cref{tab:arenas} is a starting budget and must be tuned to the
real exported models, since an undersized arena makes \texttt{AllocateTensors()} fail and that stage
report itself absent.

\begin{table}[ht]
  \centering
  \caption{Shared tensor arena (initial budget).}
  \label{tab:arenas}
  \begin{tabularx}{\textwidth}{@{}llX@{}}
    \toprule
    \textbf{Arena} & \textbf{Size} & \textbf{Notes} \\
    \midrule
    Shared (one model resident) & 48\,KiB & Sized to the audio stage (1-D conv over 8000 points); the
      motion model loads into the same arena and fits. The two flip in and out on demand, so only one
      model's tensors occupy RAM at a time \\
    \bottomrule
  \end{tabularx}
\end{table}

Sharing one arena rather than two is a deliberate choice on the memory-bound collar: it saves the
second arena outright. Enabling inference initially pushed RAM use to roughly 94\,\%; eliminating the
float-window scratch buffer (quantising directly from int16) and sharing the single arena brought it
back into the mid-80s, leaving headroom for the BLE stack and the raised main stack
(\cref{sec:firmware-collar}).
