\section{Training pipeline}
\label{sec:training-pipeline}

Meowtion learns new behaviours through a closed loop the owner drives entirely from the dashboard,
with no local toolchain: capture, label, train, deliver, run. The loop turns the collar from a
fixed-function sensor into a platform for recognising any behaviour the owner can demonstrate.

\subsection{Closed loop}

\begin{enumerate}[leftmargin=1.6em,itemsep=2pt]
  \item \textbf{Define behaviours.} The owner lists the actions to recognise (eat, drink, purr,
        scratch, litter-tray, and so on) in the dev console. This set drives both the clip-label
        dropdown and the trained model's outputs.
  \item \textbf{Capture.} With a station in training mode, paired audio and time-aligned motion are
        recorded while a registered collar is in range and uploaded to Cloud Storage (Flow~2,
        \cref{sec:architecture}).
  \item \textbf{Label.} The owner reviews clips in the dev console, sets each clip's action, trims
        dead space from the waveform, and discards poor examples.
  \item \textbf{Train.} The \texttt{train} Cloud Function trains the models server-side from the
        labelled clips and writes the results back: the quantised \texttt{.tflite} blobs to Cloud
        Storage, and the new version and label list to the Realtime Database.
  \item \textbf{Deliver.} Each station watches a model-version key; when it changes the station
        downloads the new \texttt{.tflite} files and pushes them to the collar over the air via the
        BEGIN/DATA/END/CRC sequence (\cref{sec:protocols}).
  \item \textbf{Run.} In production mode the collar classifies with the new model and the dashboard
        shows live, named behaviours. Observing those results, the owner captures more clips or adds
        behaviours and retrains, closing the loop: the collar's own data improves the model it runs
        (\cref{fig:training-loop}).
\end{enumerate}

\begin{figure}[ht]\centering
  \resizebox{0.66\textwidth}{!}{%
  \begin{tikzpicture}[every node/.style={font=\footnotesize}]
    \def\R{3.6}
    \node[mwflow] (define)  at (90:\R)   {\textbf{1.} Define\\behaviours};
    \node[mwflow] (capture) at (30:\R)   {\textbf{2.} Capture};
    \node[mwflow] (label)   at (-30:\R)  {\textbf{3.} Label};
    \node[mwflow] (train)   at (-90:\R)  {\textbf{4.} Train};
    \node[mwflow] (deliver) at (-150:\R) {\textbf{5.} Deliver\\{\scriptsize(OTA push)}};
    \node[mwflow] (run)     at (150:\R)  {\textbf{6.} Run};
    \node[align=center, text=mwmut, font=\itshape] at (0,0)
      {owner-driven loop\\[2pt]the collar supplies the data\\\emph{and} runs the model};
    \draw[mwedge] (define)  -- (capture);
    \draw[mwedge] (capture) -- (label);
    \draw[mwedge] (label)   -- (train);
    \draw[mwedge] (train)   -- (deliver);
    \draw[mwedge] (deliver) -- (run);
    \draw[mwedge] (run) -- node[midway, above left=-2pt, text=mwlavd, font=\bfseries\scriptsize]{iterate} (define);
  \end{tikzpicture}}
  \caption{The training pipeline closes into a loop. The same collar that captures the labelled data
           (steps 2--3) also runs the model trained from it (step 6); observing the live results, the
           owner refines or adds behaviours and retrains, so each pass improves the next. The loop is
           owner-driven (human-in-the-loop), not autonomous.}
  \label{fig:training-loop}
\end{figure}

\Cref{fig:capture-flow} shows how the training \emph{data} moves through the system during the capture
and label steps; \cref{fig:training-flow} (below) shows how the trained \emph{model} moves back out.

\begin{figure}[ht]\centering
  \resizebox{\textwidth}{!}{%
  \begin{tikzpicture}[node distance=12mm and 18mm, every node/.style={font=\footnotesize}]
    \node[mwnode] (collar) {Collar};
    \node[mwnode, right=of collar] (station) {Station};
    \node[mwnode, right=of station] (upl) {\texttt{upload\_clip()}};
    \node[mwnode, fill=mwlav!8, right=of upl] (store) {Storage\\\scriptsize training clips};
    \node[mwnode, fill=mwlav!8, below=15mm of station] (rtdb) {Realtime DB\\\scriptsize clip metadata};
    \node[mwnode, below=15mm of upl] (dev) {Dev console\\\scriptsize label $\cdot$ playback};
    \draw[mwedge] (collar) -- node[above,text=mwmut,font=\scriptsize]{BLE clip frames} (station);
    \draw[mwedge] (station) -- node[above,text=mwmut,font=\scriptsize]{POST (token)} (upl);
    \draw[mwedge] (upl) -- node[above,text=mwmut,font=\scriptsize]{write (admin)} (store);
    \draw[mwedge] (station) -- node[left,text=mwmut,font=\scriptsize]{metadata} (rtdb);
    \draw[mwedge] (rtdb) -- node[above,text=mwmut,font=\scriptsize]{read} (dev);
    \draw[mwedge] (store) -- node[right,text=mwmut,font=\scriptsize]{audio + IMU} (dev);
  \end{tikzpicture}}
  \caption{Training-data transport. The collar streams paired audio\,+\,IMU clip frames to the station,
           which slices them into clips and uploads the bytes through \texttt{upload\_clip} to Storage
           and the clip metadata to the database; the dev console reads both to play back, label, and
           train on them.}
  \label{fig:capture-flow}
\end{figure}

\subsection{The trainer}

Training runs in the \texttt{train} Python Cloud Function, currently gated to developer accounts to
keep cloud costs down during development rather than as an inherent limit (\cref{sec:cloud-backend}).
It reads the account's labelled clips (metadata from the Realtime
Database, audio and motion bytes from Cloud Storage) and derives the class set directly from the
labels present, so the network's outputs follow whatever behaviours the owner defined. It then trains
the two cascade stages independently:

\begin{itemize}[leftmargin=1.5em]
  \item a small 1-D convolutional network over the motion windows (six IMU channels), and
  \item a larger 1-D convolutional network over the 8\,kHz audio windows.
\end{itemize}

Each input is windowed with overlap and normalised exactly as the collar normalises it at inference,
so the on-device and training transforms match (\cref{sec:on-device-ai}). The trained Keras models are
converted to TensorFlow Lite with full int8 post-training quantisation using a representative dataset,
so the exported \texttt{.tflite} weights and activations are int8 and the input and output tensors
carry the scale and zero-point the collar uses to quantise its own inputs. \Cref{fig:training-flow}
shows the whole path, from the training trigger to the model running on the collar.

\begin{figure}[ht]
  \centering
  \begin{tikzpicture}[node distance=12mm, every node/.style={font=\footnotesize}]
    \node[mwterm] (dev) {Dev console: ``Train''};
    \node[mwflow, below=of dev, text width=6cm, align=left] (train)
      {\texttt{train()}: reads labelled clips (RTDB labels + Storage \texttt{wav}/\texttt{imu}); trains int8 IMU + audio models};
    \node[mwnode, fill=mwlav!8, below=15mm of train, xshift=-30mm, text width=3cm] (store)
      {Storage\\\texttt{models/<uid>/\allowbreak *.tflite}};
    \node[mwnode, fill=mwlav!8, below=15mm of train, xshift=30mm, text width=3cm] (rtdb)
      {Realtime DB\\\texttt{models}: version + labels};
    \node[mwflow, below=38mm of train] (station) {Station};
    \node[mwterm, below=12mm of station] (collar) {Collar: verify CRC, load model};
    \draw[mwedge] (dev) -- node[right,text=mwmut]{POST (dev token)} (train);
    \draw[mwedge] (train) -- (store);
    \draw[mwedge] (train) -- (rtdb);
    \draw[mwedge] (store) -- node[left,text=mwmut,pos=0.4]{download} (station);
    \draw[mwedge] (rtdb) -- node[right,text=mwmut,pos=0.4]{version bump} (station);
    \draw[mwedge] (station) -- node[right,text=mwmut]{OTA (BLE, CRC)} (collar);
  \end{tikzpicture}
  \caption{Training and over-the-air model delivery. \texttt{train()} reads the labelled clips, writes
           the quantised models to Storage and the new version and labels to the database, and the
           station detects the version bump, downloads the models, and pushes them to the collar.}
  \label{fig:training-flow}
\end{figure}

\subsection{Label-agnostic by design}

The trainer is not hard-coded to a fixed set of actions; it learns whatever label set the owner
defines. This makes Meowtion a platform rather than a fixed-function gadget: the same pipeline that
recognises eating and drinking can be pointed at any distinguishable behaviour the owner can capture
and label. The class set flows end to end as data, from the dashboard's action list, through the
trainer, to the index-based on-collar classifier, with names attached only at the edges.

\subsection{Delivery contract}

Training outputs are written to predictable locations so the station finds them without coordination:
the model blobs land in Cloud Storage as \texttt{models/<uid>/\allowbreak imu\_model.tflite} and
\texttt{models/<uid>/\allowbreak audio\_model.tflite}, the new version is mirrored into each of the owner's
stations' configuration so their version gate fires, and the label list is stored in
\texttt{users/<uid>/models} so the dashboard can map a class index back to a name across label-set
changes (\cref{sec:cloud-backend}).
