\section{Cloud backend}
\label{sec:cloud-backend}

The backend is Firebase~\cite{firebase}, Google's hosted backend platform (sign-in, database, file
storage, and serverless functions). It provides owner authentication, a Realtime Database (RTDB)
for live state and history, Cloud Storage for training clips and model blobs, and Python Cloud
Functions (2nd generation) for server-side logic. There is no custom server to operate. The whole
backend lives in \texttt{app/firebase/} and deploys with \texttt{firebase deploy} from that folder.

\subsection{Authentication}

Owners sign in with Firebase Authentication (email/password or Google). The collar and station do
\emph{not} authenticate as users: a station carries a scoped per-device token, and a collar carries
nothing (the station relays it). The credential surface on battery devices is thus a single revocable
token (\cref{sec:security-privacy}).

\subsection{Backend data flow}

\Cref{fig:backend-flow} shows how the tiers exchange data in normal operation. The dashboard reads the
RTDB, both a small per-cat live read and a cached history read (\cref{sec:dashboard}). The station
writes telemetry and events to the RTDB and uploads training clips to Cloud Storage through the
\texttt{upload\_clip} function. The station also relays the collar and delivers models to it over BLE.
Clients never write Cloud Storage directly (\cref{sec:security-privacy}). The training path that
produces those models appears separately in \cref{fig:training-flow}.

\begin{figure}[ht]
  \centering
  \resizebox{0.92\textwidth}{!}{%
  \begin{tikzpicture}[node distance=13mm and 32mm, every node/.style={font=\footnotesize}]
    \node[mwnode] (dash) {Dashboard};
    \node[mwnode, below=14mm of dash] (station) {Station};
    \node[mwnode, below=14mm of station] (collar) {Collar};
    \node[mwnode, fill=mwlav!8, right=of dash] (rtdb) {Realtime Database\\\scriptsize current, events, models, config};
    \node[mwnode, fill=mwlav!8, right=of station] (store) {Cloud Storage\\\scriptsize clips, models};
    \draw[mwedge] (dash) -- node[above,text=mwmut]{read (live + history)} (rtdb);
    \draw[mwedge] (station) -- node[above,pos=0.55,text=mwmut]{telemetry, events} (rtdb);
    \draw[mwedge] (station) -- node[below,text=mwmut]{clips (via \texttt{upload\_clip}); download models} (store);
    \draw[mwedge] (collar) -- node[left,text=mwmut]{BLE relay / OTA} (station);
  \end{tikzpicture}}
  \caption{Backend data flow in normal operation. The dashboard reads the database; the station writes
           live state and events and uploads clips; the collar is relayed and updated through the
           station.}
  \label{fig:backend-flow}
\end{figure}

\subsection{Realtime Database schema}

State and history live in the RTDB, keyed by owner so the security rules scope every read and write to
the owning account. \Cref{tab:db} lists the main paths.

\begin{table}[ht]
  \centering
  \caption{Realtime Database paths.}
  \label{tab:db}
  \begin{tabularx}{\textwidth}{@{}lX@{}}
    \toprule
    \textbf{Path} & \textbf{Contents} \\
    \midrule
    \texttt{users/<uid>/devices/<token>}        & A registered station (\texttt{type:station}) with its
      relayed \texttt{cats/<catId>} subtree, or a registered collar entry \\
    \texttt{.../cats/<catId>/current}           & The cat's live state (version, class, confidence, steps, battery) \\
    \texttt{.../cats/<catId>/events}            & Discrete activity events (carry model version + class index) \\
    \texttt{.../devices/<station>/weather}      & Current and recent local weather for the station's location \\
    \texttt{users/<uid>/models}                 & Current model version, status, and the label list \\
    \texttt{users/<uid>/actions}                & The behaviours to recognise (managed from the dev console) \\
    \texttt{deviceTokens/<token>/owner}         & Maps a device token to its owning account \\
    \texttt{config/demoOwner}                   & UID whose data backs the public read-only demo \\
    \texttt{config/devAccounts/<uid>}           & Accounts allowed to use the dev/training tools \\
    \bottomrule
  \end{tabularx}
\end{table}

Events store the model version and class index rather than a pre-resolved name. The dashboard maps the
index to a label using that version's label list, so historical events stay correct as the owner
changes their behaviour set.

\subsection{Cloud Storage}

Cloud Storage holds the training clips (audio plus aligned motion) and the trained model blobs in an
owner-scoped layout: clips at \texttt{training/<owner>/\allowbreak <collar>/<ts>.\{wav,imu\}} and models
at \texttt{models/<uid>/\allowbreak \{imu,audio\}\_model.tflite}. Training clips are private to the owner
and \emph{not} world-readable, even in the demo, because home audio must never be public. Model blobs
are public-read so the token-only station can fetch one to push to the collar, and are written only by
the trainer.

\subsection{Cloud Functions}

Two functions provide the only server-side logic. Both run on the function's ambient service account,
so no service-account key is in the repository.

\begin{description}[leftmargin=1.6em,style=nextline]
  \item[\texttt{train}] Server-side model training, \texttt{POST} with a developer account's Firebase
        ID token and gated additionally by \texttt{config/devAccounts/<uid>}. It reads the account's
        labelled clips, trains the int8 IMU and audio models (\cref{sec:training-pipeline}), uploads
        them to \texttt{models/<uid>/}, updates \texttt{users/<uid>/models}, and mirrors the new
        version into each of the owner's station configs to trigger delivery.
  \item[\texttt{upload\_clip}] Authenticated clip upload for the token-only station. The station has
        no login, so it cannot satisfy the Cloud Storage rules directly. It POSTs the clip bytes here
        with its device token; the function verifies \texttt{deviceTokens/<token>/owner} and writes the
        file as administrator into that owner's area. The rules can then deny all direct client writes
        while still admitting an authorised station.
\end{description}

\subsection{Demo data simulator}

So the public demo always presents a populated, believable dashboard, a scheduled Cloud Function
maintains a fully simulated companion collar (``Purrminator'') on the demo account. It is standalone:
no station, always online. On first run it backfills roughly six months of realistic activity history,
weighted by time of day to match an average cat (crepuscular, busy around dawn and dusk, resting
overnight), with matching daily weather from a free historical feed. Thereafter it wakes on a short
schedule and commits each activity as an event when that action ends, so new events arrive at varying
intervals. The simulator writes only the demo account's own data and never touches a real collar.

\subsection{Weather context}

Local weather lets the dashboard read a habit change against its environment, since a warm spell that
raises drinking and suppresses appetite should not be mistaken for illness. The station polls a keyless
forecast feed (Open-Meteo) for its provisioned location and writes current and recent weather under its
device record. The demo simulator backfills weather from the same provider's historical archive
alongside the simulated activity.

\subsection{Configuration trust boundary}

The \texttt{config} subtree is readable by clients but not writable by them, and is maintained from the
Firebase console. The system can thus publish facts the clients need (which account backs the demo,
which accounts are developers) without letting a client grant itself privileges.
