\section{Dashboard}
\label{sec:dashboard}

The owner-facing application is a Streamlit dashboard backed by a small set of static pages for
the parts that need the Firebase web SDK directly (sign-in, account and device management,
the developer/training console, and the privacy policy). It is hosted on Streamlit Community
Cloud and deploys automatically from the repository's main branch.

\subsection{Why a hybrid}

Authentication and device pairing are easiest with the Firebase web SDK in the browser, while the
data views and charts are easiest in Streamlit's Python. The two halves are served from the same
origin so they can share a login: the static sign-in page writes the Firebase ID token into a
same-origin cookie, and the Python dashboard reads that cookie to identify the owner. A separate
flag cookie marks the public read-only demo.

\subsection{Pages and views}

\begin{table}[ht]
  \centering
  \caption{Dashboard surfaces.}
  \label{tab:dash}
  \begin{tabularx}{\textwidth}{@{}lX@{}}
    \toprule
    \textbf{Surface} & \textbf{Role} \\
    \midrule
    \texttt{app.py} (Streamlit)   & Live cat state, activity history, charts, account controls \\
    \texttt{account.html}         & Sign-in, account, device pairing over Web Serial \\
    \texttt{dev.html}             & Developer/training console: capture, label, train, mode \\
    \texttt{privacy.html}         & Privacy policy \\
    \bottomrule
  \end{tabularx}
\end{table}

\subsection{Live view and history}

The dashboard reads the cat's live ``current'' record and renders the latest behaviour, and it
flattens the event history into a table for charts (totals per behaviour, frequency, and trends
over time). Activity charts are rendered with a transparent background so they sit on the page's
brand canvas rather than in a clashing panel.

\subsection{Read-only demo}

A public demo lets anyone explore the product without an account. It reads the demo owner's data
(permitted by the security rules) but every write control and function trigger is disabled, and
the developer/training console is available in a read-only variant. The demo is genuinely
read-only because it is enforced by the backend rules, not just hidden in the UI: the demo
account is world-readable but write-locked (\cref{sec:security-privacy}).
