\section{Communication protocols}
\label{sec:protocols}

Four contracts tie the tiers together: the BLE telemetry the collar advertises, the audio clip-frame
format it streams during training, the over-the-air model protocol, and the cloud HTTP it is all
relayed over. The two BLE wire formats are defined once in shared headers
(\texttt{firmware/common/meow\_protocol.h} and \texttt{meow\_ota.h}) that \emph{both} firmwares
include, so the collar (Zephyr) and station (ESP-IDF) builds cannot drift out of sync.

\subsection{Telemetry advertisement}

In production mode the collar publishes its result in the manufacturer-specific field of its BLE
advertisement, so the station can read it passively without connecting. The packet is eight bytes
(\cref{tab:telemetry}); the collar's BLE address identifies which collar it is, so no identifier is
carried in the payload. The \texttt{version} byte selects how the two result bytes are read: version~1
is the simulated behaviour state machine, version~2 is real on-device classification.

\begin{table}[ht]
  \centering
  \caption{Eight-byte telemetry packet (manufacturer AD data).}
  \label{tab:telemetry}
  \begin{tabularx}{\textwidth}{@{}llX@{}}
    \toprule
    \textbf{Bytes} & \textbf{Field} & \textbf{Meaning} \\
    \midrule
    0--1 & company id   & \texttt{0xFFFF} (test/development identifier) \\
    2    & version      & \texttt{1} = simulated state, \texttt{2} = real on-device classification \\
    3    & state\,/\,class & v1: state (0 sleep \dots 5 groom); v2: predicted class index (\texttt{0xFF} = unknown) \\
    4    & activity\,/\,confidence & v1: activity level 0--100; v2: confidence 0--100 \\
    5--6 & steps        & step count, \texttt{uint16}, little-endian \\
    7    & battery      & battery level 0--100 \\
    \bottomrule
  \end{tabularx}
\end{table}

Because the class is an index, not a name, the station and dashboard resolve it through the label list
for the running model version, which keeps history correct across label-set changes
(\cref{sec:cloud-backend}).

\subsection{Audio clip-streaming frame}

While a station is subscribed in training mode, the collar streams continuously on the audio
characteristic in 100\,ms frames. Each frame is a packed header followed by an audio payload and a
time-aligned IMU payload; the station concatenates frames and slices them into fixed-length training
clips.

\begin{lstlisting}[language=C,caption={Clip-frame header (\texttt{firmware/common/meow\_protocol.h}).},label={lst:clip}]
#define MEOW_CLIP_MAGIC 0x574F454Du   /* 'MEOW' little-endian, marks each frame start */

struct __attribute__((packed)) meow_clip_hdr {
    uint32_t magic;
    uint8_t  version;      /* 1 = 16-bit PCM audio, 2 = 8-bit G.711 mu-law */
    uint8_t  imu_axes;     /* 6 = accel x/y/z + gyro x/y/z */
    uint16_t imu_rate_hz;
    uint32_t audio_bytes;  /* audio payload size (one 100 ms frame) */
    uint32_t imu_bytes;    /* IMU payload size */
    uint32_t audio_rate;   /* output sample rate, for the station's WAV header */
};
/* frame layout:  [ meow_clip_hdr ][ audio payload ][ IMU payload ] */
\end{lstlisting}

To halve the data on the BLE link the collar encodes audio as 8-bit G.711 $\mu$-law (\texttt{version}
2); the station decodes it back to 16-bit PCM for a normal WAV. The $\mu$-law codec is defined inline
in the same shared header, so the encode on the collar and the decode on the station are guaranteed to
match, and this is the same representation the model is trained and run on (\cref{sec:on-device-ai}).

\subsection{GATT services}

The collar exposes two custom 128-bit GATT services that share a common Meowtion base UUID: the
\emph{audio} service used for clip streaming, and the \emph{OTA} service used for model delivery. The
OTA service uses three UUIDs distinguished by a \texttt{0x0b}\,$\ast\ast$ group, one each for its
control, data, and status characteristics. Both firmwares derive these UUIDs from the same header, the
collar through Zephyr's encoding macro and the station from the raw little-endian byte arrays the
header also provides.

\subsection{Over-the-air model protocol}

Model delivery is a small framed protocol over the OTA service. The client (station) writes opcodes to
the \emph{control} characteristic, streams the model to the \emph{data} characteristic, and the collar
acknowledges on a \emph{status} notification. A transfer is: \texttt{BEGIN} (carrying the target slot,
the total length, and a CRC), a stream of write-with-response \texttt{DATA} chunks (each chunk's write
response is the flash-write flow control), then \texttt{END}. The collar erases the slot on
\texttt{BEGIN}, writes chunks as they arrive, and on \texttt{END} re-reads the model from flash,
verifies the CRC against the announced value, and reports the outcome.

\begin{lstlisting}[language=C,caption={OTA control opcodes, BEGIN payload, and status codes (\texttt{firmware/common/meow\_ota.h}).},label={lst:ota}]
/* CONTROL characteristic: first byte is an opcode; BEGIN is followed by meow_ota_begin. */
enum { MEOW_OTA_OP_BEGIN = 1, MEOW_OTA_OP_END = 2, MEOW_OTA_OP_ABORT = 3 };

struct __attribute__((packed)) meow_ota_begin {
    uint8_t  slot;       /* 0 = IMU model, 1 = audio model */
    uint32_t total_len;  /* model size in bytes */
    uint32_t crc32;      /* CRC-32/IEEE over the model blob */
};

/* STATUS characteristic (notify): the collar's acknowledgement. */
enum { MEOW_OTA_ST_READY = 1, MEOW_OTA_ST_RECEIVING = 2, MEOW_OTA_ST_OK = 3,
       MEOW_OTA_ST_ERR_SIZE = 5, MEOW_OTA_ST_ERR_CRC = 7, MEOW_OTA_ST_ERR_SEQ = 8 };
\end{lstlisting}

Crucially, \texttt{END} verifies and stages the model but does \emph{not} swap the live model from
inside the BLE callback. Committing a model is stack-heavy, and doing it in the radio callback
crash-looped the device; the actual swap is deferred to a service call on the main thread's larger
stack (\cref{sec:firmware-collar}). That split, together with the raised main stack, is what made OTA
reliable on hardware. A persisted slot magic marks a slot that holds a valid committed model, so the
collar can tell an empty slot from a written one across a reboot.

\subsection{Cloud transport}

Above BLE, the station speaks ordinary HTTPS to Firebase: it writes telemetry to the Realtime Database
through its REST interface, authorised by its device token, and POSTs training-clip bytes to the
\texttt{upload\_clip} Cloud Function with that same token (\cref{sec:cloud-backend}). The dashboard is
a separate client of the same database. No bespoke transport or message broker is involved; the
contract at this tier is just the database schema and the function's request format.
