\section{Security and privacy}
\label{sec:security-privacy}

Privacy is a design property of Meowtion, not a setting bolted on afterwards. The strongest guarantees
come from where work happens (on the collar) and how data is scoped (per owner, by enforced rules).

\subsection{Audio never leaves the collar}

The microphone is used only for on-device classification. Audio is processed on the collar and
discarded; it is never written to flash in normal operation and never transmitted in production mode.
The collar's uplink is a classification result (a class index and a confidence), not a sensor stream.
The one exception is deliberate and owner-driven: in training mode the owner chooses to capture clips
to teach a behaviour, and those clips are stored privately to their own account and are never
world-readable, even for the public demo.

\subsection{Per-owner isolation}

All of an owner's data lives under \texttt{users/<uid>/}, and the database and storage rules scope
every read and write to the owning account, so owners (and their token-authenticated devices) touch
only their own data. Isolation is enforced by the backend, not the client. The Storage rules go
further and deny \emph{all} direct client writes: training clips are written only by the
\texttt{upload\_clip} function and are readable only by their owning account, and model blobs are
public-read (so the token-only station can fetch one) but writable only by the trainer.

\subsection{Devices hold tokens, not credentials}

A station authenticates with a scoped, revocable per-device token rather than the owner's password, and
a collar holds nothing at all. A lost or compromised device therefore cannot expose the owner's
credentials, and it can be revoked simply by deleting its token from the account, with no password
change and no effect on the owner's other devices. \Cref{fig:token} traces the token's lifecycle.

\begin{figure}[ht]\centering
  \begin{tikzpicture}[node distance=10mm and 14mm, every node/.style={font=\footnotesize}]
    \node[mwterm] (mint) {Dashboard mints token T\\\scriptsize records \texttt{deviceTokens/T/owner}};
    \node[mwflow, below=of mint, text width=6.6cm, align=left] (store)
      {Station stores T in NVS and presents it on every cloud request (no login)};
    \node[mwflow, below=15mm of store, xshift=-26mm, text width=4cm, align=left] (rules)
      {Database rules: T maps to the owner, so writes to that owner's subtree are allowed};
    \node[mwflow, below=15mm of store, xshift=26mm, text width=4cm, align=left] (upl)
      {\texttt{upload\_clip} verifies T's owner and writes the clip as admin};
    \node[mwterm, below=38mm of store] (revoke) {Revoke: delete the token; all its access is denied};
    \draw[mwedge] (mint) -- (store);
    \draw[mwedge] (store) -- (rules);
    \draw[mwedge] (store) -- (upl);
    \draw[mwedge] (rules) -- (revoke);
    \draw[mwedge] (upl) -- (revoke);
  \end{tikzpicture}
  \caption{Credential-token lifecycle. A device is minted a scoped token that maps to its owner; the
           token authorises database writes and clip uploads, and deleting it revokes the device.}
  \label{fig:token}
\end{figure}

\subsection{The demo is read-only by enforcement}

The public demo reads a single designated demo account's data, which the rules make world-readable, but
that account is write-locked at the backend: the demo cannot change any data or trigger any function.
The read-only behaviour is enforced by the rules, not merely hidden in the interface, so it holds even
against a hand-crafted client.

\subsection{What is and is not a secret}

The Firebase web API key shipped in the static front end is a public client identifier, not a secret;
security is enforced by Authentication and the rules, not by hiding the key. No service-account key is
present in the repository, and the functions run on their ambient service account. Device Wi-Fi
credentials and firmware secrets are excluded by \texttt{.gitignore}. Everything required to understand
and build the system is tracked; only true secrets are not.

\subsection{Owner data rights}

The account page lets an owner export everything stored about their cats and delete their account,
devices, and data. This supports the access and erasure expectations of data-protection regimes such as
the UK GDPR. The bundled privacy policy is a starting point and is marked for legal review before any
public launch.
