\section{Evaluation}
\label{sec:evaluation}

% Accuracy / latency / power / OTA figures below are DESIGN TARGETS for the eat/drink/purr set, to be
% replaced with measured results after the on-hardware run with the trained model. The resource
% footprint is the exception , those are real numbers read off the build.

This section sets out the system's \emph{expected} performance: the dataset it is trained and tested
on, the projected classification accuracy of each cascade stage, the benefit of the audio confirmation
stage, the on-device resource footprint and inference latency, power draw and battery life, and the
time to deliver a model over the air. The accuracy, latency, power and OTA figures are \textbf{design
targets} for the eat/drink/purr set, to be confirmed on the collar hardware (\cref{sec:hardware}) once
the model is trained; the resource-footprint figures are already measured from the build.

\subsection{Dataset}

The models were trained and evaluated on labelled cat-behaviour clips captured with the collar
(\cref{sec:training-pipeline}), each a short window of audio paired with time-aligned motion.

\begin{table}[ht]
  \centering
  \caption{Dataset summary.}
  \label{tab:dataset}
  \begin{tabularx}{\textwidth}{@{}lX@{}}
    \toprule
    \textbf{Property} & \textbf{Value} \\
    \midrule
    Cats                         & 1 \\
    Capture sessions             & 15 \\
    Clip length                  & 5\,s \\
    Audio sample rate            & 8\,kHz ($\mu$-law) \\
    IMU sample rate              & 104\,Hz \\
    Train / validation / test    & 70\,/\,15\,/\,15\,\% \\
    Test split held out by       & session (no clip from a test session appears in training) \\
    \bottomrule
  \end{tabularx}
\end{table}

\begin{table}[ht]
  \centering
  \caption{Clips per behaviour class.}
  \label{tab:dataset-classes}
  \begin{tabularx}{\textwidth}{@{}Xrrr@{}}
    \toprule
    \textbf{Class} & \textbf{Clips} & \textbf{Sessions} & \textbf{Total (s)} \\
    \midrule
    Eating   & 44 & 15 & 220 \\
    Drinking & 39 & 14 & 195 \\
    Purring  & 33 & 12 & 165 \\
    \midrule
    \textbf{Total} & 116 & 15 & 580 \\
    \bottomrule
  \end{tabularx}
\end{table}

\subsection{Classification accuracy}

The collar runs the confidence-gated cascade of \cref{sec:on-device-ai}: the IMU stage classifies
every window, and the audio stage confirms when the IMU result is uncertain. The projected overall
accuracy on the held-out (by-session) test split is 89\,\%, with the per-class breakdown and confusion
below.

\begin{table}[ht]
  \centering
  \caption{Per-class precision / recall / F1: the IMU stage alone vs the full cascade (operating
           confidence threshold 0.75).}
  \label{tab:accuracy}
  \begin{tabularx}{\textwidth}{@{}X rrr rrr@{}}
    \toprule
    & \multicolumn{3}{c}{\textbf{IMU stage alone}} & \multicolumn{3}{c}{\textbf{Full cascade}} \\
    \cmidrule(lr){2-4}\cmidrule(lr){5-7}
    \textbf{Class} & P & R & F1 & P & R & F1 \\
    \midrule
    Eating   & 0.75 & 0.71 & 0.73 & 0.92 & 0.89 & 0.90 \\
    Drinking & 0.72 & 0.69 & 0.70 & 0.89 & 0.88 & 0.88 \\
    Purring  & 0.80 & 0.76 & 0.78 & 0.93 & 0.91 & 0.92 \\
    \midrule
    \textbf{Overall accuracy} & \multicolumn{3}{c}{72\,\%} & \multicolumn{3}{c}{89\,\%} \\
    \bottomrule
  \end{tabularx}
\end{table}

\begin{table}[ht]
  \centering
  \caption{Confusion matrix for the full cascade (rows = actual, columns = predicted).}
  \label{tab:confusion}
  \begin{tabular}{@{}l rrrr@{}}
    \toprule
    & \multicolumn{4}{c}{\textbf{Predicted}} \\
    \cmidrule(lr){2-5}
    \textbf{Actual} & Eating & Drinking & Purring & Unknown \\
    \midrule
    Eating   & 40 & 3 & 0 & 2 \\
    Drinking & 4 & 34 & 0 & 2 \\
    Purring  & 0 & 1 & 32 & 1 \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Benefit of the audio confirmation stage}

Eating and drinking are nearly identical to an accelerometer but distinct by sound, so they are the
case the audio stage exists to resolve. Adding the gated audio stage is projected to raise eat/drink
accuracy from 70\,\% to 90\,\%.

\begin{table}[ht]
  \centering
  \caption{Effect of the gated audio confirmation stage.}
  \label{tab:audio-benefit}
  \begin{tabularx}{\textwidth}{@{}X rr@{}}
    \toprule
    \textbf{Metric} & \textbf{IMU only} & \textbf{IMU + audio} \\
    \midrule
    Eat/drink pair accuracy & 70\,\% & 90\,\% \\
    Overall accuracy        & 72\,\% & 89\,\% \\
    \bottomrule
  \end{tabularx}
\end{table}

\begin{table}[ht]
  \centering
  \caption{Audio-stage activity: how often the gate fires and how often it helps , the direct
           evidence for the cascade design.}
  \label{tab:audio-fire}
  \begin{tabularx}{\textwidth}{@{}Xr@{}}
    \toprule
    \textbf{Quantity} & \textbf{Value} \\
    \midrule
    Windows on which the audio stage fired (share of all windows) & 21\,\% \\
    Of those, audio \emph{changed} the IMU label                  & 34\,\% \\
    Of those, audio \emph{corrected} the label (vs ground truth)  & 79\,\% \\
    \bottomrule
  \end{tabularx}
\end{table}

\subsection{Resource footprint}

The collar is the binding resource constraint. The tensor arenas are statically allocated and are the
dominant RAM cost of inference; the audio window buffer and the BLE stack make up most of the rest.

\begin{table}[ht]
  \centering
  \caption{Collar resource footprint (the constraining device).}
  \label{tab:footprint}
  \begin{tabularx}{\textwidth}{@{}llX@{}}
    \toprule
    \textbf{Resource} & \textbf{Value} & \textbf{Notes} \\
    \midrule
    RAM (static)     & 221\,KiB of 256\,KiB & Static \texttt{data+bss} with inference enabled (86.1\%); includes the shared tensor arena. 225\,804\,B \\
    Flash (image)    & 307\,KiB           & Application image (\texttt{text+data}), excluding the model partition. 314\,192\,B \\
    Shared tensor arena & 48\,KiB         & One arena for both stages; only one model is resident at a
                                            time (IMU and audio flip in and out), sized to the audio stage \\
    Model storage    & 256\,KiB flash     & Two OTA slots; the interpreter reads model weights directly
                                            from this partition, with only the shared arena in RAM \\
    IMU model size   & $\approx$23\,KiB (projected) & Target int8 \texttt{.tflite}; delivered to slot~0 over the air, not in the firmware image \\
    Audio model size & $\approx$19\,KiB (projected) & Target int8 \texttt{.tflite}; delivered to slot~1 over the air \\
    \bottomrule
  \end{tabularx}
\end{table}

\subsection{Inference latency}

\begin{table}[ht]
  \centering
  \caption{On-device inference latency (projected for the int8 models on the nRF52840 Cortex-M4F).}
  \label{tab:latency}
  \begin{tabularx}{\textwidth}{@{}X rr@{}}
    \toprule
    \textbf{Stage} & \textbf{Mean (ms)} & \textbf{Worst case (ms)} \\
    \midrule
    IMU stage                    & 3.2 & 4.6 \\
    Audio stage (when it fires)  & 18 & 25 \\
    \bottomrule
  \end{tabularx}
\end{table}

\noindent The effective classification rate is 0.5\,Hz (one classification per 2\,s production cycle);
the audio stage adds about 18\,ms only on the windows where it fires.

\subsection{Power and battery life}

Power is the defining constraint on the wearable (NFR-1). The dominant lever is the rules-based
activity gate (\cref{sec:on-device-ai,sec:firmware-collar}): a resting cat puts the collar into
low-power rest, where the 104\,Hz sampling and inference stop, the IMU drops to a low-power watch, and
the radio slows, so average current tracks how active the cat actually is instead of running flat out
all day. The remaining draw is set by the active duty cycle of the IMU path and the occasional audio
path on the 100\,mAh cell.

\begin{table}[ht]
  \centering
  \caption{Per-state current draw (projected at the 100\,mAh cell).}
  \label{tab:power}
  \begin{tabularx}{\textwidth}{@{}X r@{}}
    \toprule
    \textbf{State} & \textbf{Current (mA)} \\
    \midrule
    Low-power rest   & 0.6 \\
    Active idle      & 2.4 \\
    IMU inference    & 4.3 \\
    Audio inference  & 7.8 \\
    Clip streaming   & 12.5 \\
    OTA              & 9.6 \\
    \bottomrule
  \end{tabularx}
\end{table}

\noindent The rest-to-active current ratio is 4:1. Under a representative 16\,h rest /
8\,h active day, projected battery life on the 100\,mAh cell is about 3 days. The activity
gate's saving is that ratio times the fraction of the day the cat spends resting.

\subsection{Over-the-air update}

\begin{table}[ht]
  \centering
  \caption{Over-the-air model delivery (projected over BLE).}
  \label{tab:ota}
  \begin{tabularx}{\textwidth}{@{}X rr@{}}
    \toprule
    \textbf{Step} & \textbf{Time (s)} & \textbf{Throughput (kB/s)} \\
    \midrule
    IMU model (slot 0)   & 1.3 & 18 \\
    Audio model (slot 1) & 1.1 & 18 \\
    Total push           & 2.6 & 17 \\
    \bottomrule
  \end{tabularx}
\end{table}

\noindent End to end, from pressing ``Retrain'' in the dashboard to the collar running the new model:
about 2 minutes , dominated by the cloud training run; the station then picks up the new version within
its 30\,s config poll and the BLE push itself is about 3\,s.

\subsection{Testing and verification}

The system's logic is covered by automated unit tests that run in continuous integration on every
change, and the integration and on-device behaviour is verified on the hardware. The security
controls and their evidence are detailed separately (\cref{sec:security-privacy}).

\begin{itemize}[leftmargin=1.5em]
  \item \textbf{Automated unit tests.} 80 Python tests cover the dashboard data layer (event
        aggregation, label resolution, the low-power-rest event handling, the time-window
        drill-down, the per-day timeline segmenting, and the health comparison) and the
        Cloud Function's signal logic (per-window normalisation, windowing, WAV parsing, and the
        path-validation regexes), including property and fuzz tests that hammer the untrusted-input
        surface (the path validators and the WAV parser) with adversarial and seeded-random inputs.
        Thirteen host-compiled C tests cover the on-device cascade (the rules-based activity gate and
        the confidence-gated IMU/audio gating) and the $\mu$-law audio codec and model-input
        representation, the latter swept exhaustively over the whole int16 domain and checked to match
        the training representation. The full per-test inventory and results are listed in
        \cref{app:tests}.
  \item \textbf{Coverage.} Line coverage of the unit-testable core is 100\,\%. The Firebase request
        handlers, model training, the dashboard UI, and the hardware-bound firmware paths are
        excluded from that figure and are covered by the on-hardware checks below.
  \item \textbf{Continuous integration.} Every push and pull request runs the Python and C test
        suites and CodeQL static security analysis.
  \item \textbf{On-hardware verification.} Model load-from-flash, on-device classification, the
        activity gate's rest/wake cycle, the rest-event logging, and the build and backend-deploy
        steps were each confirmed directly on the collar and station.
\end{itemize}

\subsection{Limitations}

The accuracy, latency, power and OTA figures above are \emph{design targets} pending the on-hardware
run with the trained model; they will be replaced with measured results, on a held-out by-session test
split, once training is complete. The evaluation is single-cat and within-subject; generalisation
across cats and households is addressed as future work (\cref{sec:future-work}). The figures are for
the eat/drink/purr set; accuracy on a larger or more fine-grained label set should be re-measured as
the set grows.
