\section{Future work}
\label{sec:future-work}

Meowtion already runs the whole loop on real hardware: capture, label, train, deliver over the air,
classify on the collar, and surface the result live. What follows is the roadmap from that working
system toward its real goal (noticing a cat's health change early), ordered so the near-term items
that finish the product come first and the research-scale ones last.

\begin{description}[leftmargin=1.6em,style=nextline]
  \item[From activity to health insight] The dashboard already tracks each habit and flags a large,
        sustained change against the cat's own baseline (\cref{sec:dashboard}). The next step is the
        health analytics the product is really about: baselines that adapt as the cat ages, anomaly
        detection over a multi-day trend rather than single events, and a concise, vet-shareable
        summary when a habit drifts. This is the shortest path from an activity tracker to a genuine
        early-warning tool, and it needs no new hardware.
  \item[Richer, per-cat recognition] Extend the labelled set beyond eat/drink/purr to scratching,
        litter-tray use, grooming, and play, and add a short per-cat fine-tune pass so recognition
        sharpens for each animal's own motion and sound. Both fall straight out of the label-agnostic
        pipeline (\cref{sec:training-pipeline}).
  \item[Longer battery life] Add true hardware wake-on-motion (an IMU interrupt waking the SoC from
        System OFF) and tune the activity gate's thresholds and the audio path's duty cycle against
        the measured power budget (\cref{sec:evaluation}) for the lowest possible rest current.
  \item[Multiple cats and shared stations] Several cats are already supported, each on its own collar
        with per-cat data and a dashboard switcher. The open problems are \emph{model} generalisation
        (training on one cat and performing well on another, since the evaluation is currently
        single-cat and within-subject) and telling apart two cats that share a bowl or station.
  \item[Full firmware over the air] Today only models are delivered wirelessly; extend the same
        staged, CRC-checked mechanism (\cref{sec:protocols}) to complete firmware updates, so a
        deployed collar never needs a cable.
  \item[Field validation and publication] Move the accuracy study beyond a single cat with a larger,
        multi-household dataset, ideally paired with vet-confirmed events to test the early-warning
        claim directly, then write up the confidence-gated cascade as a preprint and a venue paper.
\end{description}
