\section{Data-model reference}
\label{sec:appendix-schema}

A consolidated reference for the cloud data model introduced in \cref{sec:cloud-backend}: the Realtime
Database tree, the Cloud Storage layout, and the request format of the HTTP Cloud Functions.

\subsection{Realtime Database tree}

\begin{lstlisting}[numbers=none,caption={Realtime Database layout (owner-scoped).},label={lst:rtdb}]
users/<uid>/
  devices/
    <stationToken>/            type: "station"
      cats/<catId>/
        current                { ver, cls, conf, steps, battery, ts }
        events/<startMs>        { start, durationSec, type | cls, ver }
      weather/                  { current, history }
      config/                   { modelVer, mode, captureForce, ... }
      seen/<collarId>           unregistered collars heard nearby
    <collarId>/                 type: "collar"  (registry entry; name, simulated)
  models/                       { version, status, labels[], results }
  actions/                      behaviours to recognise (dev console)
  profile/                      { consent, location }
deviceTokens/<token>/owner      maps a device token to its owning account
config/
  demoOwner                     uid backing the public read-only demo
  devAccounts/<uid>             accounts allowed to use the dev/training tools
\end{lstlisting}

Events store the model \texttt{version} and class \texttt{cls} index rather than a resolved name; the
dashboard maps the index to a label through that version's \texttt{labels} list, so history stays
correct across label-set changes. Low-power rest spans from the activity gate are logged the same way
but carry \texttt{type:\,"rest"} instead of a model class, so the timeline accounts for the resting
period rather than leaving a gap (\cref{sec:firmware-collar}).

\subsection{Cloud Storage layout}

\begin{lstlisting}[numbers=none,caption={Cloud Storage layout.},label={lst:storage}]
training/<owner>/<collar>/<ts>.wav    training-clip audio   (private to owner; no client writes)
training/<owner>/<collar>/<ts>.imu    time-aligned motion   (private to owner)
models/<uid>/imu_model.tflite         trained IMU model     (public-read; written by train only)
models/<uid>/audio_model.tflite       trained audio model   (public-read; written by train only)
\end{lstlisting}

\subsection{Cloud Function requests}

\begin{description}[leftmargin=1.6em,style=nextline]
  \item[\texttt{train}] \texttt{POST} with header \texttt{Authorization: Bearer <Firebase ID token>}
        of a developer account. Trains from the account's labelled clips; returns the new version and
        label set. Gated additionally by \texttt{config/devAccounts/<uid>}.
  \item[\texttt{upload\_clip}] \texttt{POST} the raw clip bytes with header
        \texttt{Authorization: Bearer <device token>} (kept out of the URL/query string). The function
        resolves \texttt{deviceTokens/<token>/owner} and writes the file as administrator into that
        owner's Storage area. No Firebase login is involved.
  \item[\texttt{simulate\_now}] \texttt{POST} with header \texttt{Authorization: Bearer <Firebase ID
        token>} of a developer account; manually kicks a demo-simulator run (the simulator otherwise
        runs on a per-minute schedule). Gated by \texttt{config/devAccounts/<uid>}.
\end{description}
