\section{Automated test inventory}
\label{app:tests}

% Identifiers/literals contain underscores; \detokenize renders them literally without manual escaping
% and avoids the \path clash with TikZ used elsewhere in this document.
\providecommand{\tn}[1]{\texttt{\detokenize{#1}}}

% A filled-in test-case form (IEEE-829 style). One per automated test.
% #1 ID  #2 Title  #3 Component  #4 Type  #5 Preconditions  #6 Input/data  #7 Procedure  #8 Expected  #9 Status
\newcommand{\testcase}[9]{%
  \par\smallskip\noindent
  \begingroup\footnotesize
  \arrayrulecolor{mwline}\setlength{\arrayrulewidth}{0.4pt}\renewcommand{\arraystretch}{1.25}%
  \begin{tabularx}{\textwidth}{|>{\bfseries\color{mwink}}p{3.0cm}|X|}
    \hline
    \rowcolor{mwlav!10}\multicolumn{2}{|@{\hspace{6pt}}l|}{%
      {\bfseries\color{mwlavd}Test case #1}\hfill{\color{mwmut}#4}}\\
    \hline
    Title & #2\\\hline
    Component & #3\\\hline
    Preconditions & #5\\\hline
    Input / test data & #6\\\hline
    Procedure & #7\\\hline
    Expected result & #8\\\hline
    Actual result & As expected\\\hline
    Status & \textbf{#9}\\\hline
  \end{tabularx}%
  \endgroup\smallskip
}

This appendix is the test-case specification for the project's automated tests: one filled form per
test, recording its scenario, inputs, expected result and status. The suites run in continuous
integration on every push and pull request (\cref{sec:evaluation}). As of the run on 2026-07-01,
\textbf{all 93 tests pass}: \textbf{80} Python tests (run with \texttt{pytest}) and \textbf{13}
host-compiled C tests for the on-device cascade and the audio codec. Line coverage of the unit-testable core
(\tn{meowtion_dash/data.py} and the Cloud Function's \tn{main.py}) is \textbf{100\,\%}. Parametrised
tests are recorded as a single form whose input field lists every case. ``Actual result: as expected''
denotes the assertions held when the suite last ran.

\subsection{Firmware tests (C, host-compiled)}

Compiled and run on the host from \tn{firmware/test} (no Zephyr, no hardware). The classifier tests
link strong overrides of the model hooks (declared \texttt{weak} in \tn{classifier.c}) and count
audio-stage invocations, so every gating branch is checked.

\testcase{ACT-01}{Motion keeps the collar awake}{Activity gate , \tn{activity.c}}{Unit test (host C)}
{\tn{activity_reset()} called; step \tn{dt}=2\,s.}
{100 consecutive readings at peak motion \tn{ACT_STILL_MG}+50 (above the stillness threshold).}
{Call \tn{activity_should_rest(peak, 2)} for each reading.}
{Every call returns ``do not rest'' , motion never drops the collar to low-power rest.}{Pass}

\testcase{ACT-02}{Sustained stillness trips rest}{Activity gate , \tn{activity.c}}{Unit test (host C)}
{\tn{activity_reset()} called; step \tn{dt}=2\,s.}
{Still readings (peak 0) every 2\,s, spanning the hold-off \tn{ACT_REST_HOLDOFF_S}.}
{Call \tn{activity_should_rest(0, 2)} repeatedly across the hold-off.}
{Returns ``no rest'' for every step before the hold-off, then ``rest'' once stillness is sustained past it.}{Pass}

\testcase{ACT-03}{Motion resets the still-timer}{Activity gate , \tn{activity.c}}{Unit test (host C)}
{\tn{activity_reset()} called; still-timer advanced to just before the hold-off.}
{One reading at the threshold \tn{ACT_STILL_MG}, then a still reading.}
{Call \tn{activity_should_rest(ACT_STILL_MG, 2)} then \tn{activity_should_rest(0, 2)}.}
{Does not rest immediately , the motion reset the timer, so a full hold-off is required again.}{Pass}

\testcase{CLF-01}{Confident IMU skips audio}{Cascade , \tn{classifier.c}}{Unit test (host C)}
{\tn{conf_threshold}=0.75; audio confirmation enabled; audio model present.}
{IMU = (class 2, confidence 0.90); audio = (5, 0.99).}
{Call \tn{clf_classify()} for the cycle; count audio invocations.}
{Audio stage not run (0 calls); result class = 2 (the confident IMU result).}{Pass}

\testcase{CLF-02}{Unsure IMU consults audio}{Cascade , \tn{classifier.c}}{Unit test (host C)}
{\tn{conf_threshold}=0.75; audio confirmation enabled; audio model present.}
{IMU = (2, 0.40) below threshold; audio = (5, 0.95).}
{Call \tn{clf_classify()} for the cycle; count audio invocations.}
{Audio run exactly once; result class = 5 (the more-confident audio result).}{Pass}

\testcase{CLF-03}{Audio kept only when more confident}{Cascade , \tn{classifier.c}}{Unit test (host C)}
{\tn{conf_threshold}=0.75; audio confirmation enabled; audio model present.}
{IMU = (2, 0.40); audio = (5, 0.30) , less confident than the IMU.}
{Call \tn{clf_classify()} for the cycle; count audio invocations.}
{Audio run once but discarded; result class = 2 (the IMU result is kept).}{Pass}

\testcase{CLF-04}{Disabled confirmation never runs audio}{Cascade , \tn{classifier.c}}{Unit test (host C)}
{\tn{conf_threshold}=0.75; audio confirmation \emph{disabled}; audio model present.}
{IMU = (2, 0.40) below threshold; audio = (5, 0.95).}
{Call \tn{clf_classify()} for the cycle; count audio invocations.}
{Audio stage never runs (0 calls); result class = 2.}{Pass}

\testcase{CLF-05}{No audio buffer skips audio}{Cascade , \tn{classifier.c}}{Unit test (host C)}
{\tn{conf_threshold}=0.75; audio confirmation enabled; no PCM buffer this cycle.}
{IMU = (2, 0.40) below threshold; PCM = NULL (the production IMU-only path).}
{Call \tn{clf_classify(NULL, NULL, 0)}; count audio invocations.}
{Audio stage skipped (0 calls); result class = 2.}{Pass}

\testcase{CLF-06}{No audio model skips audio}{Cascade , \tn{classifier.c}}{Unit test (host C)}
{\tn{conf_threshold}=0.75; audio confirmation enabled; no audio model present.}
{IMU = (2, 0.40) below threshold; audio model absent.}
{Call \tn{clf_classify()} for the cycle; count audio invocations.}
{Audio stage skipped (0 calls); result class = 2.}{Pass}

\subsection{Audio codec and representation tests (C, host-compiled)}

From \tn{firmware/test/test_audio_codec.c}, over the header-only \texttt{G.711} µ-law codec
(\tn{meow_protocol.h}) and the model-input transform (\tn{audio_codec.h}). The codec invariants are
swept \emph{exhaustively} over the whole int16 domain (only 65\,536 values); the transform is fuzzed
over seeded-random windows. Writing them caught and fixed a codec bug, pinned in ACO-02 below.

\testcase{ACO-01}{Codec byte round-trip is consistent}{µ-law codec , \tn{meow_protocol.h}}{Property/fuzz (host C)}
{None.}{All 256 µ-law bytes.}
{For each byte \tn{b} whose \tn{ulaw_decode(b)} is non-zero, check \tn{ulaw_encode(ulaw_decode(b))} equals \tn{b}.}
{Holds for every byte. The two zero codes \texttt{0x7F} (-0) and \texttt{0xFF} (+0) alias to \texttt{+0} on re-encode and are exempt (standard G.711).}{Pass}

\testcase{ACO-02}{Codec preserves the sign of a non-zero sample}{µ-law codec , \tn{meow_protocol.h}}{Property/fuzz (host C)}
{None.}{Every int16 of magnitude $\geq 256$ (exhaustive).}
{Check \tn{ulaw_decode(ulaw_encode(s))} keeps the sign of \tn{s} and is non-zero.}
{Holds for all. Writing this caught \tn{INT16_MIN} ($-32768$) encoding to near-silence , negating it as int16 overflowed back to $-32768$; fixed by taking the magnitude in a wider \tn{int}.}{Pass}

\testcase{ACO-03}{Model-input transform: length and per-sample round-trip}{Audio representation , \tn{audio_codec.h}}{Property/fuzz (host C)}
{None.}{4\,000 seeded-random 16\,kHz windows and output caps.}
{Call \tn{audio_to_model_pcm}; check the length is \texttt{floor(in/2)} clamped to the cap and each output equals the µ-law round-trip of the pairwise-averaged pair.}
{Holds for every window.}{Pass}

\testcase{ACO-04}{Training wire representation equals inference representation}{Audio representation , \tn{audio_codec.h}}{Differential (host C)}
{None.}{4\,000 seeded-random 16\,kHz windows.}
{Compare \tn{audio_to_model_pcm} (inference) against the training path (decimate 16$\to$8\,kHz, \tn{ulaw_encode}, then the station's \tn{ulaw_decode}).}
{Identical samples , the collar trains and infers on the same representation, so there is no train/inference mismatch.}{Pass}

\subsection{Cloud Function tests (Python)}

From \tn{app/firebase/functions/tests}; the Firebase SDKs are stubbed in \tn{conftest.py}.
\tn{test_dsp.py} pins the exact representation the model trains on (and the collar reproduces at
inference); the \tn{test_helpers.py} path-validation forms are security tests.

\testcase{DSP-01}{Normalisation is zero-mean and bounded}{DSP , \tn{main.py}}{Unit test (pytest)}
{Firebase SDKs stubbed.}{One channel \texttt{[1,2,3]}.}{Call \tn{per_window_norm(x)}.}
{Output mean is 0 to within \texttt{1e-5}; all values lie within \texttt{[-1,1]}.}{Pass}

\testcase{DSP-02}{Normalisation matches the reference formula}{DSP , \tn{main.py}}{Unit test (pytest)}
{Firebase SDKs stubbed.}{\texttt{[0,10]}.}{Call \tn{per_window_norm(x)}.}
{Output equals \texttt{[-1,1]} exactly (mean 5 removed, then divided by max-abs 5).}{Pass}

\testcase{DSP-03}{Channels normalised independently}{DSP , \tn{main.py}}{Unit test (pytest)}
{Firebase SDKs stubbed.}{Two channels \texttt{[[0,100],[4,104]]}.}{Call \tn{per_window_norm(x)}.}
{Each channel's mean is 0 independently (channels are not pooled).}{Pass}

\testcase{DSP-04}{Short input padded to one window}{DSP , \tn{main.py}}{Unit test (pytest)}
{Firebase SDKs stubbed.}{3$\times$2 array; \tn{win}=5, \tn{hop}=2.}{Call \tn{windows(arr, win, hop)}.}
{Exactly one window of shape \texttt{(5,2)}; the last two rows are zero-padded.}{Pass}

\testcase{DSP-05}{Window slides with the hop}{DSP , \tn{main.py}}{Unit test (pytest)}
{Firebase SDKs stubbed.}{\tn{arange(10)} reshaped \texttt{(10,1)}; \tn{win}=4, \tn{hop}=3.}{Call \tn{windows(arr, win, hop)}.}
{Three windows starting at indices 0,\,3,\,6; each shape \texttt{(4,1)}.}{Pass}

\testcase{DSP-06}{Dataset windows IMU clips}{DSP , \tn{main.py}}{Unit test (pytest)}
{\tn{LABELS}=[\tn{eat},\tn{drink}].}{One IMU clip labelled \tn{drink}.}{Call \tn{build_dataset([clip], "imu")}.}
{\tn{X} shape \texttt{(1, IMU\_WIN, IMU\_AXES)}; \tn{y}=[1] (the label index).}{Pass}

\testcase{DSP-07}{Dataset windows audio clips}{DSP , \tn{main.py}}{Unit test (pytest)}
{\tn{LABELS}=[\tn{eat},\tn{drink}].}{One audio clip labelled \tn{eat}.}{Call \tn{build_dataset([clip], "audio")}.}
{\tn{X} shape \texttt{(1, AUDIO\_WIN, 1)}; \tn{y}=[0].}{Pass}

\testcase{DSP-08}{Empty IMU clip skipped}{DSP , \tn{main.py}}{Unit test (pytest)}
{\tn{LABELS}=[\tn{eat}].}{A clip with a zero-length IMU array.}{Call \tn{build_dataset([clip], "imu")}.}
{No windows produced (\tn{len(X)}=0) , empty clips never reach training.}{Pass}

\testcase{VAL-01}{Valid ids accepted}{Path validation (security) , \tn{main.py}}{Unit test (pytest)}
{Firebase SDKs stubbed.}{4 inputs: \tn{cat_64582d}, \tn{abc-123_XYZ}, \tn{A}, \tn{0}.}
{Match each against \tn{_SAFE_ID}.}{All four are accepted.}{Pass}

\testcase{VAL-02}{Traversal and separators rejected}{Path validation (security) , \tn{main.py}}{Unit test (pytest)}
{Firebase SDKs stubbed.}{10 inputs: \tn{../etc}, \tn{a/b}, \tn{a..b/}, \tn{a.b}, \tn{a b}, \tn{a;b}, empty, \tn{a/../b}, \tn{..}, \texttt{a\textbackslash b}.}
{Match each against \tn{_SAFE_ID}.}{All ten are rejected (no path can escape the owner's prefix).}{Pass}

\testcase{VAL-03}{Numeric timestamp accepted}{Path validation (security) , \tn{main.py}}{Unit test (pytest)}
{Firebase SDKs stubbed.}{\tn{1719600000000} (all digits).}{Match against \tn{_SAFE_TS}.}
{Accepted.}{Pass}

\testcase{VAL-04}{Non-numeric timestamps rejected}{Path validation (security) , \tn{main.py}}{Unit test (pytest)}
{Firebase SDKs stubbed.}{7 inputs: \tn{12a}, \tn{-1}, \tn{1.0}, \tn{../1}, empty, \tn{1 2}, \tn{0x10}.}
{Match each against \tn{_SAFE_TS}.}{All seven are rejected.}{Pass}

\testcase{WAV-01}{WAV parser round-trips samples}{WAV parsing , \tn{main.py}}{Unit test (pytest)}
{Firebase SDKs stubbed.}{An 8\,kHz mono 16-bit WAV of \texttt{[0,1,-1,32767,-32768]}.}{Call \tn{parse_wav(buf)}.}
{Returns rate 8000 and exactly those samples (including the int16 extremes).}{Pass}

\testcase{WAV-02}{Sample rate read from header}{WAV parsing , \tn{main.py}}{Unit test (pytest)}
{Firebase SDKs stubbed.}{A 16\,kHz WAV.}{Call \tn{parse_wav(buf)}.}
{Parsed rate = 16000 (read from the \texttt{fmt} chunk, not assumed).}{Pass}

\testcase{WAV-03}{Non-WAV input rejected}{WAV parsing , \tn{main.py}}{Unit test (pytest)}
{Firebase SDKs stubbed.}{Non-RIFF bytes.}{Call \tn{parse_wav(buf)}.}
{Raises \tn{ValueError}.}{Pass}

\testcase{WAV-04}{Missing data chunk rejected}{WAV parsing , \tn{main.py}}{Unit test (pytest)}
{Firebase SDKs stubbed.}{A WAV with a \texttt{fmt} chunk but no \texttt{data} chunk.}{Call \tn{parse_wav(buf)}.}
{Raises \tn{ValueError}.}{Pass}

\subsection{Property and fuzz tests (Python)}

From \tn{app/firebase/functions/tests/test_fuzz.py}, over the untrusted-input surface , the path
validators and the WAV parser, which take fully attacker-controlled values. Each form encodes an
\emph{invariant} and throws both targeted adversarial inputs and tens of thousands of seeded-random
inputs at it (stdlib \tn{random}; no external fuzzing engine, so they run in ordinary CI). Writing
them surfaced and fixed two issues, both pinned below: the id/timestamp validators accepted a trailing
newline (Python's \texttt{\$} also matches just before a newline, so the end is now anchored with
\tn{\Z}); and \tn{parse_wav} raised a raw \tn{struct.error} on a truncated \texttt{fmt} chunk instead
of \tn{ValueError}.

\testcase{FUZ-01}{Trailing terminators rejected by the id validator}{Path validation (security) , \tn{main.py}}{Property/fuzz (pytest)}
{Firebase SDKs stubbed.}{Valid ids suffixed with a trailing newline, carriage return, or NUL byte (7 cases).}
{Match each against \tn{_SAFE_ID}.}{All rejected: the end is anchored with \tn{\Z}, not \texttt{\$}, so a trailing newline cannot ride into the Storage path.}{Pass}

\testcase{FUZ-02}{Trailing terminators rejected by the timestamp validator}{Path validation (security) , \tn{main.py}}{Property/fuzz (pytest)}
{Firebase SDKs stubbed.}{Numeric timestamps suffixed with a trailing newline, CR, or NUL (4 cases).}
{Match each against \tn{_SAFE_TS}.}{All rejected.}{Pass}

\testcase{FUZ-03}{Accepted ids contain only the allowed charset}{Path validation (security) , \tn{main.py}}{Property/fuzz (pytest)}
{Firebase SDKs stubbed.}{40\,000 seeded-random strings from the allowed set plus tricky characters (newline, CR, NUL, dot, slash, backslash, space).}
{For every string \tn{_SAFE_ID} accepts, check each character is in \texttt{[A-Za-z0-9\_-]}.}
{No accepted string contains a disallowed character.}{Pass}

\testcase{FUZ-04}{Accepted timestamps contain only digits}{Path validation (security) , \tn{main.py}}{Property/fuzz (pytest)}
{Firebase SDKs stubbed.}{40\,000 seeded-random strings from digits plus tricky and letter characters.}
{For every string \tn{_SAFE_TS} accepts, check each character is a digit.}
{No accepted string contains a non-digit.}{Pass}

\testcase{FUZ-05}{Truncated fmt chunk fails cleanly}{WAV parsing , \tn{main.py}}{Property/fuzz (pytest)}
{Firebase SDKs stubbed.}{A WAV whose \texttt{fmt} chunk header declares a body the buffer does not contain.}
{Call \tn{parse_wav(buf)}.}{Raises \tn{ValueError}, not a raw \tn{struct.error} (the sample-rate read is bounds-checked).}{Pass}

\testcase{FUZ-06}{Parser survives arbitrary bytes}{WAV parsing , \tn{main.py}}{Property/fuzz (pytest)}
{Firebase SDKs stubbed.}{6\,000 seeded-random byte buffers, half prefixed with a valid RIFF/WAVE header.}
{Call \tn{parse_wav} on each.}{Every call returns \texttt{(int16 samples, rate)} or raises \tn{ValueError} , never another exception, never a hang.}{Pass}

\testcase{FUZ-07}{Parser survives adversarial chunk sizes}{WAV parsing , \tn{main.py}}{Property/fuzz (pytest)}
{Firebase SDKs stubbed.}{6\,000 syntactically-plausible chunk streams with adversarial sizes (zero, odd, and up to $2^{31}$).}
{Call \tn{parse_wav} on each.}{Returns cleanly or raises \tn{ValueError} , the size arithmetic never overruns a read.}{Pass}

\testcase{FUZ-08}{Normalisation stays finite and bounded under fuzz}{DSP , \tn{main.py}}{Property/fuzz (pytest)}
{Firebase SDKs stubbed.}{3\,000 random int16 arrays of varying shape.}
{Call \tn{per_window_norm(x)}.}{Output is always finite (no NaN/inf) and bounded to about \texttt{[-1, 1]}.}{Pass}

\testcase{FUZ-09}{Windows are always exactly one window long}{DSP , \tn{main.py}}{Property/fuzz (pytest)}
{Firebase SDKs stubbed.}{3\,000 random \texttt{(length, win, hop)} combinations.}
{Iterate \tn{windows(arr, win, hop)}.}{Every yielded window has exactly \tn{win} rows (short input is zero-padded, never short).}{Pass}

\subsection{Dashboard data-layer tests (Python)}

From \tn{app/dashboard/tests}, over the data layer that turns raw Firebase JSON into the chart-ready
frame: event parsing, the time-window drill-down, the per-day segmenting behind the timeline, and the
health-watch comparison.

\testcase{DAT-01}{Duration formatting}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{\tn{fmt_dur(45)} and \tn{fmt_dur(125)}.}{Call \tn{fmt_dur}.}
{Returns \tn{45s} and \tn{2m 05s} respectively.}{Pass}

\testcase{DAT-02}{Class index maps to label}{Data layer , \tn{data.py}}{Unit test (pytest)}
{\tn{labels}=[\tn{eat},\tn{drink}].}{A \texttt{ver==2} event with \tn{cls}=0 and \texttt{durationSec}=120.}{Call \tn{activity_dataframe}.}
{Activity is \tn{Eating} (label 0, canonicalised); \tn{event_duration}=2.0\,min.}{Pass}

\testcase{DAT-03}{Rest sentinel uses type, not class}{Data layer , \tn{data.py}}{Unit test (pytest)}
{\tn{labels}=[\tn{eat},\tn{drink}].}{A \texttt{ver==2} event with \tn{cls}=254 (\texttt{0xFE}), \tn{type}=rest, \texttt{durationSec}=600.}{Call \tn{activity_dataframe}.}
{Activity is \tn{Resting} from the \tn{type} (254 is not indexed into \tn{labels}); 10.0\,min.}{Pass}

\testcase{DAT-04}{Legacy v1 event uses type}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{A v1 event (no \tn{cls}) with \tn{type}=groom.}{Call \tn{activity_dataframe}.}
{Activity is \tn{Grooming}.}{Pass}

\testcase{DAT-05}{Malformed event skipped}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{An event with no \tn{start} timestamp.}{Call \tn{activity_dataframe}.}
{The row is dropped; the resulting frame is empty (no exception).}{Pass}

\testcase{DAT-06}{Activity filter, single and list}{Data layer , \tn{data.py}}{Unit test (pytest)}
{Frame with activities \{\tn{Eat},\tn{Drink},\tn{Rest}\}.}{Filter by \tn{Eat}, then by [\tn{Eat},\tn{Rest}].}{Call \tn{filter_by_activity}.}
{Returns \{\tn{Eat}\}, then \{\tn{Eat},\tn{Rest}\}.}{Pass}

\testcase{DAT-07}{Over-time sums duration}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{Two same-day events of 120\,s and 60\,s.}{Call \tn{over_time}.}
{Per-activity total is 3.0\,min.}{Pass}

\testcase{DSH-01}{Parses nested and standalone cats}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{A device tree with a station-relayed cat and a standalone collar.}{Call \tn{iter_cats}.}
{Both are returned with name and relay resolved; the standalone's \tn{via}=\tn{None}.}{Pass}

\testcase{DSH-02}{Cat list de-duplicated}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{A mixed device tree.}{Call \tn{list_cats}.}
{Returns de-duplicated \texttt{(name, id)} pairs.}{Pass}

\testcase{DSH-03}{Day options newest-first}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{Dates 27,\,28,\,28 Jun.}{Call \tn{period_options(df, "Day")}.}
{Options \texttt{[2026-06-28, 2026-06-27]} , newest-first and de-duplicated.}{Pass}

\testcase{DSH-04}{Day and month windows}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{Mixed dates; Day=2026-06-28, Month=2026-06.}{Call \tn{filter_to_window}.}
{Day keeps 2 rows; Month keeps the June rows \{27,\,28\}.}{Pass}

\testcase{DSH-05}{Recent average without baseline}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{3 days of eating \texttt{[10,20,30]}\,min.}{Call \tn{health_signals}.}
{\tn{recent}=20.0; \tn{baseline}=\tn{None}; \tn{change_pct}=\tn{None} (too little history).}{Pass}

\testcase{DSH-06}{Time formatting}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{An epoch-milliseconds value.}{Call \tn{fmt_time}.}
{Renders a ``dd Mon $\cdot$ HH:MM'' string.}{Pass}

\testcase{DSH-07}{Weekday filter}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{A set \{Saturday, Sunday\}, and a single \tn{Monday}.}{Call \tn{filter_by_weekday}.}
{Keeps the matching weekday rows in both cases.}{Pass}

\testcase{DSH-08}{Date-range filter}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{A bounded range, and a start-only (open-ended) range.}{Call \tn{filter_by_date_range}.}
{Bounded keeps \{27,\,28\}; start-only keeps \{28\}.}{Pass}

\testcase{DSH-09}{Last N calendar days}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{Dates 26,\,27,\,28; \tn{n}=2.}{Call \tn{last_n_days(df, n=2)}.}
{Keeps the two most recent days \{27,\,28\}.}{Pass}

\testcase{DSH-10}{Habit drop versus baseline}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{Eating \texttt{[100,8,8,8,999]} over 5 days.}{Call \tn{health_signals}.}
{Latest still-accruing day (999) dropped; \tn{recent}=8.0; \tn{baseline}=100.0; \tn{change_pct}=-92.0.}{Pass}

\testcase{DSH-11}{Model labels with guards}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{\tn{models/labels} present, then \texttt{\{\}}, then \tn{None}.}{Call \tn{model_labels}.}
{Returns \texttt{[eat,drink]}, then \texttt{[]}, then \texttt{[]} (no exception).}{Pass}

\testcase{DSH-12}{Non-dict entries skipped}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{A non-dict device entry and a non-dict cat entry.}{Call \tn{iter_cats}.}
{Both are skipped; only the valid cat is returned.}{Pass}

\testcase{DSH-13}{Empty input and Week/Month options}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{Empty input; then dates spanning two months.}{Call \tn{period_options}.}
{Empty input yields \texttt{[]}; Week/Month produce labelled options (\tn{Week of ...}, \tn{July 2026}/\tn{June 2026}).}{Pass}

\testcase{DSH-14}{Empty input and Week window}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{Empty input; then a Week window from 2026-06-29.}{Call \tn{filter_to_window}.}
{Empty input handled; the Week window keeps \{2026-06-29\}.}{Pass}

\testcase{DSH-15}{Empty input and hourly path}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{Empty input; then a single-day (hourly) aggregation.}{Call \tn{over_time}.}
{Empty input yields an empty frame; the Day path sums to 2.0.}{Pass}

\testcase{DSH-16}{Empty-input guards}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{Empty input.}{Call \tn{last_n_days} and \tn{health_signals}.}
{Both return empty results (\texttt{[]} / empty frame) without raising.}{Pass}

\testcase{DSH-17}{Midnight-crossing event split per day}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{Rest at 22:00 lasting 240\,min (crosses midnight).}{Call \tn{daily_segments}.}
{Two rows: 22:00 to 24:00 on day 1 and 00:00 to 02:00 on day 2, mapped onto the shared time-of-day reference.}{Pass}

\testcase{DSH-18}{Zero-duration and empty skipped}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{A zero-duration event; then empty input.}{Call \tn{daily_segments}.}
{No rows are produced in either case.}{Pass}

\testcase{DSH-19}{Sparse edge days trimmed}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{3 days of 10 / 600 / 10\,min.}{Call \tn{trim_sparse_edge_days}.}
{The thin first and last days are trimmed; the busy middle day is kept.}{Pass}

\testcase{DSH-20}{Sparse interior day kept}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{3 days of 600 / 10 / 600\,min.}{Call \tn{trim_sparse_edge_days}.}
{The thin interior day is kept (all three remain) , only edges are trimmed.}{Pass}

\testcase{DSH-21}{Single-day and empty handled}{Data layer , \tn{data.py}}{Unit test (pytest)}
{None.}{A single-day frame; then an empty frame.}{Call \tn{trim_sparse_edge_days}.}
{The single-day frame is returned unchanged; the empty frame stays empty.}{Pass}
